% ── 2. Assumptions and Constraints ────────────────────────────────────────────
\section{Assumptions and Constraints}

Every architectural decision rests on a set of assumptions about the operating environment, the user population, and the technical context. Stating these assumptions explicitly serves two purposes: it makes the design defensible by showing which conditions it was optimized for, and it identifies the boundary conditions under which the design would need to be revisited. The assumptions below are organized by domain.

\subsection{User and Scale Assumptions}

The initial target population is individual consumer households in a single geographic market. The platform is expected to onboard thousands of households in its first three years, with growth concentrated in the first and second years following product launch. Each household is treated as an independent deployment unit with its own account, device registry, and access permissions. Household size is assumed to be small — typically one to five members — and at most one member is designated the primary account holder with full administrative access. The platform does not need to support enterprise or multi-site deployments within the scope of this design.

Peak concurrent usage is assumed to follow a residential usage pattern, concentrated in early morning and evening hours, with a meaningful fraction of devices generating telemetry continuously throughout the day regardless of active user sessions. The platform is designed to handle a device fleet that is one to two orders of magnitude larger than the active user session count at any given moment, reflecting the fact that devices remain connected and reporting even when no consumer is actively using the application.

\subsection{Device and Connectivity Assumptions}

All smart-home modules are assumed to connect to the cloud platform over the household's existing broadband internet connection via WiFi. The platform does not provision or manage the home network itself; it assumes a functioning internet connection is already present. Devices are assumed to maintain persistent connections to AWS IoT Core rather than polling, which allows the platform to push commands to devices without requiring devices to initiate every exchange. Connection quality is assumed to be adequate for reliable MQTT messaging; intermittent connectivity is handled at the protocol level through message queuing and acknowledgment, but the platform does not need to account for extended offline periods lasting days.

The software communication protocol between the cloud platform and the device firmware is within the scope of this architecture to specify. The hardware design and low-level firmware implementation are handled by a separate engineering team. This architecture assumes that devices can be updated over the air and that the firmware team will implement the device-side of the specified protocol. Device identity is assumed to be provisioned at the factory, with each unit receiving a unique certificate and thing identity before it leaves manufacturing.

\subsection{Data and Storage Assumptions}

Telemetry retention is assumed to be bounded: the platform retains granular time-series data for a rolling window of 90 days, after which older data may be aggregated or expired. Camera video retention is configurable per household with a maximum of 90 days. Operational metadata — device registrations, household configurations, automation rules, and user accounts — has no defined expiration and is retained indefinitely unless explicitly deleted by the consumer. The total data volume per household is expected to be modest for metadata but substantial for video-enabled households, as analyzed in the cost section.

\subsection{Security and Compliance Assumptions}

The platform is assumed to operate under standard consumer data protection expectations for a US-based product. No industry-specific regulatory frameworks (such as HIPAA or PCI-DSS) are assumed to apply, though the architecture does not preclude compliance with these frameworks if the business requires it in the future. All data in transit is encrypted using TLS. All data at rest is encrypted using AWS-managed keys. Device credentials are provisioned using X.509 certificates managed through AWS IoT Core's certificate authority. Consumer passwords are managed by Amazon Cognito and are never stored or processed by application code.

\subsection{Operational Assumptions}

The platform is operated by a small engineering and operations team typical of an early-stage product. The design therefore prioritizes managed AWS services over self-operated infrastructure, accepting a degree of vendor dependency in exchange for reduced operational complexity. Deployments are assumed to be fully automated through a CI/CD pipeline. Blue/green or canary deployment patterns are assumed for production releases to limit the blast radius of defective deployments. The platform is deployed in a single AWS region for the initial launch, with the expectation that multi-region expansion would be evaluated if the product reaches a user base that justifies geographic distribution or stricter data residency requirements.
