% ── 5. Cloud Architecture Quality Attributes ("-ilities") ────────────────────
\section{Cloud Architecture Quality Attributes}

The quality attributes of a system — often called ``-ilities'' — describe how the system behaves under varying conditions rather than what it does. For a consumer IoT platform, these attributes are not equally weighted: a system that processes commands correctly but only some of the time, or that is fast in good conditions but fragile under load, does not meet the expectations of a household customer who depends on the platform to control their home remotely. The attributes below are ordered by priority as determined by the requirements of this platform.

\subsection{Scalability}

Scalability is the system's ability to handle growth in users, devices, and data volume without architectural change. This is the primary quality concern for an early-stage platform expecting rapid adoption. The architecture addresses scalability by preferring services that scale automatically and without operator intervention. AWS Lambda scales horizontally to the number of concurrent requests by instantiating additional execution environments on demand; there are no servers to pre-provision or resize. Amazon API Gateway scales automatically with incoming request volume up to its regional limits. AWS IoT Core is a managed service with no inherent device connection ceiling that the platform operator must manage. DynamoDB in on-demand capacity mode adjusts read and write throughput automatically in response to actual traffic without requiring capacity planning. The one dimension of the architecture that requires more deliberate capacity consideration is Kinesis Video Streams, where ingestion bandwidth scales with the number of active cameras and the encoding bitrate of each stream; however, KVS itself is a managed service and its underlying capacity is not a constraint the platform operator controls.

\subsection{Availability}

% TODO: Discuss multi-AZ deployment of managed services, IoT Core SLA, API Gateway SLA, DynamoDB SLA. Lambda regional redundancy. Cognito availability.

Availability reflects the fraction of time the platform is reachable and functional for both consumers and devices. The target availability for the consumer-facing application is 99.9\% monthly uptime, which permits approximately 44 minutes of downtime per month. For device telemetry ingestion, a higher effective availability is desirable because devices generate data continuously and lost telemetry events cannot be recovered after the fact. The architecture achieves this by relying exclusively on AWS managed services, each of which carries a published service-level agreement at or above the platform's availability target. API Gateway, Lambda, DynamoDB, and AWS IoT Core are all deployed across multiple availability zones within the chosen AWS region by default, meaning that a failure in a single availability zone does not take down the service.

\subsection{Fault Tolerance}

% TODO: Discuss how Lambda retries, IoT Core message persistence, Step Functions error handling, DynamoDB replication, and S3 durability contribute to fault isolation and recovery.

Fault tolerance is the system's ability to continue operating correctly in the presence of component failures, rather than failing entirely when one part of the system becomes unavailable. The architecture is designed so that failures in one path do not cascade into failures in other paths. Device telemetry ingestion through IoT Core is independent of the user-facing API path through API Gateway and Lambda; a surge of device events does not degrade the responsiveness of the consumer dashboard, and vice versa. Step Functions provides durable workflow execution with built-in retry logic, meaning that a transient Lambda failure during an automation workflow does not silently drop the workflow — it is retried according to a configurable backoff policy and fails with a recorded error state if retries are exhausted.

\subsection{Security}

Security is a non-negotiable attribute for a platform that grants remote access to physical locks, cameras, and other sensitive home devices. The architecture enforces security at every layer. All device-to-cloud communication is authenticated using X.509 client certificates managed by AWS IoT Core, ensuring that only provisioned devices can publish or subscribe to the message broker. All consumer-to-cloud communication is authenticated through Amazon Cognito tokens validated at the API Gateway boundary before any Lambda function is invoked. AWS WAF enforces rate limiting and blocks common web attack patterns at the edge, reducing the attack surface before requests reach application logic. Data at rest in DynamoDB, S3, and Timestream is encrypted using AWS-managed keys. The architecture maintains a strict separation between consumer identity (Cognito) and device identity (IoT Core certificates) so that a compromised user credential cannot be used to impersonate a device, and a compromised device certificate cannot be used to access user account data.

\subsection{Performance}

Performance in this context refers primarily to the command round-trip latency experienced by a consumer issuing a device command — the time between the user tapping a button and receiving confirmation that the device has responded. The architecture targets a round-trip time of under two seconds for standard control operations. API Gateway and Lambda together introduce approximately 50 to 150 milliseconds of processing latency for a typical authenticated request. The IoT Core publish-subscribe model adds another 100 to 300 milliseconds for command delivery and device acknowledgment over a typical residential internet connection. DynamoDB read latency for a single-item get is typically in the single-digit millisecond range. CloudFront caches static frontend assets at edge locations close to the consumer, minimizing the time-to-interactive for the web application on initial load.

A secondary performance consideration is telemetry write throughput. The platform must absorb tens of thousands of device messages per second at fleet scale without queuing or dropping events. IoT Core and Timestream are both designed for high-throughput ingest, and the pipeline between them — IoT rules routing messages to Timestream — operates asynchronously so that the device's publish acknowledgment is not blocked by the database write.

\subsection{Cost Efficiency}

Cost efficiency is addressed in detail in Section~\ref{sec:cost}, but it is worth identifying as a quality attribute because the architectural choices that support scalability and availability also have direct cost implications. The use of serverless compute (Lambda) and on-demand database capacity (DynamoDB) means that the platform incurs costs proportional to actual usage rather than provisioned capacity, which is particularly favorable during the early growth phase when traffic is unpredictable. The separation of the camera media path from the general IoT telemetry path prevents video bandwidth from inflating the per-message cost of IoT operations. The use of CloudFront for static asset delivery reduces origin traffic to S3, lowering data transfer costs. The most significant cost inefficiency in the architecture is the per-unit-time cost of Kinesis Video Streams storage, which the platform mitigates by offering tiered retention options and an S3 archival path for long-term video retention at lower cost.
