% ── 12. Conclusion ────────────────────────────────────────────────────────────
\section{Conclusion}

The LightsPlease architecture proposes a cloud-native, managed-services-first platform for a consumer smart-home product built on Amazon Web Services. The design separates the system into three functional paths — web and API ingress, device connectivity and telemetry, and camera media — and connects these paths through a serverless application layer that handles business logic, automation orchestration, and data routing. This separation reflects the fundamental difference between user-driven and device-driven traffic patterns that characterizes IoT platforms and prevents either traffic type from degrading the other's performance.

The primary strength of the architecture is its alignment between workload characteristics and service capabilities. AWS IoT Core handles the device messaging and connection management that would be complex and fragile to self-operate. Lambda's consumption-based model keeps compute costs proportional to usage during a growth phase where traffic is unpredictable. DynamoDB's single-digit millisecond latency supports responsive user interactions. Kinesis Video Streams provides a managed camera ingest and playback path that would otherwise require custom media server infrastructure. Together, these choices allow a small engineering team to operate a platform at scale without the operational burden of managing the underlying infrastructure.

The architecture's most significant tradeoff is its deep dependency on AWS-proprietary services. This dependency is a deliberate and reasonable choice at this stage of the product's life, but it should be acknowledged as a long-term architectural constraint. The second significant tradeoff is the cost of the camera media path: KVS storage is the dominant cost driver for camera-equipped households, and the platform's unit economics are materially affected by camera adoption rate and retention policies. The cost model developed in Section~\ref{sec:cost} provides a quantitative basis for evaluating these policies and establishing default configurations that balance consumer value against platform cost.

The architecture is designed to evolve. Immediate near-term improvements include enabling DynamoDB point-in-time recovery, implementing Lambda dead-letter queues for the IoT-triggered processing path, and defining explicit RPO and RTO targets that can guide backup and failover strategy. Medium-term improvements include right-sizing DynamoDB capacity from on-demand to provisioned auto-scaling as traffic patterns stabilize, implementing per-household storage quotas to prevent camera cost overruns, and evaluating S3 Intelligent-Tiering for archival video. The long-term path to multi-region deployment via DynamoDB Global Tables and Route 53 failover is architected to be additive — the data layer and routing layer can be extended to support multiple regions without requiring a redesign of the application or device connectivity layers.

The design presented in this report demonstrates that a consumer IoT platform serving thousands of households can be built on managed AWS services in a way that is operationally simple, horizontally scalable, secure by default, and cost-transparent. The architectural decisions made are defensible given the kata's requirements, and the tradeoffs are clearly documented so that the engineering team can revisit them as the product and its operating conditions evolve.
